{\bf\Large Acknowledgement} \\ [1em] 

First, I would like to thank Prof. Dr. J체rgen Giesl for giving me the opportunity to work on an ongoing and relevant topic.
Secondly, I would like to thank apl. Prof. Dr. Thomas Noll for agreeing to be the second referee of my thesis.

Thirdly, I would like to thank Jera Hensel, who supervised me during my thesis. I want to thank her for the many patient answers she gave me no matter how obvious the solution was. She did not only answer questions, but also got proactive herself and helped me creating better results by pointing out my failures and encouraging me during the whole process. Also I want to thank her for the possibility to write the underlying program the way I wanted to without any restrictions or limits regarding the way of approaching the topic. 

Also I want to thank my girlfriend Nadine Vinkelau and all my friends, who encouraged me during my whole studies and not only accepted that I often was short on time, but also backed me up during the whole process. Especially I want to thank my good friend Tobias R채wer, who explained many topics to me over and over again throughout the whole Bachelor studies to help me pass my exams without demanding anything in return. Thanks to his selfless behaviour I got this far within only three years.

Finally I want to thank my parents for giving me the possibility to fulfil my desire to study at a worldwide known university. Without the financial support I would not have had this opportunity. \\ \\

\paragraph{Erkl채rung} Ich versichere hiermit, dass ich die vorliegende Arbeit selbstst채ndig verfasst und keine
anderen als die angegebenen Quellen und Hilfsmittel benutzt sowie Zitate kenntlich
gemacht habe.\newline \\
Aachen, den \myformat\today

\begin{tabular}{lp{2em}l} 
	\hspace{4cm} \\\cline{1-1}\cline{3-3} 
	Timo Bergerbusch
\end{tabular}
