{\bf\Large Acknowledgement} \\ [1em] 

First, I would like to thank Prof. Dr. J체rgen Giesl for giving me the opportunity to work on a GER:aktuelle topic.

Secondly I would like to thank Jera Hensel, who supervised me during my thesis. I want to thank her for the many patience answers she gave my no matter how obvious the solution was and encouraging me during the whole process. Also I want to thank her for the possibility to write the underlying program the way I wanted to without any restrictions or limits regarding the way of approaching the topic. 

Also I want to thank my girlfriend Nadine Vinkelau and all my friends, who encouraged me during my whole studies and not only accepting that I often was short on time, but also strengthen my back during the whole process. Especially I want to thank my good friend Tobias R채wer, who explained many topics to me over and over again to help me pass my exams without demanding anything in return. Thanks to his selfless behaviour I got this far within 3 years.

Finally I want to thank my parents for giving me the possibility to fullfill my desire to study at a worldwide known university. Without the financial support I would not have had this opportunity. \\ \\

\paragraph{Erkl채rung} Ich versichere hiermit, dass ich die vorliegende Arbeit selbstst채ndig verfasst und keine
anderen als die angegebenen Quellen und Hilfsmittel benutzt sowie Zitate kenntlich
gemacht habe.\newline \\
Aachen, den \today

\begin{tabular}{lp{2em}l} 
	\hspace{4cm} \\\cline{1-1}\cline{3-3} 
	Timo Bergerbusch
\end{tabular}
